Tool-driven Revolutions in Archaeological Science

### 2018-12-19 analysis/paper/jcaa-paper.Rmd
adding reference to paper by Fitzpatrick

In light of the discussion in the philosophy of science, whether paradigm-shifts are tool- or idea-driven, we offer the hypothesis that a tool-driven change of paradigms is underway in the biological sciences and, more subtly, in archaeology. The tool which drives this change is the use of scripting languages in computational approaches of these disciplines. We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence. This also is an indicator for the increase in popularity of R.

@@

As with data, sharing code in public takes a little time to ensure that it is fit for others to read. Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach. This suggests to us that there is some scope to improve the ease and efficiency of code-use in archaeology by studying the exsiting practice and providing a model that others can easily adapt to their own research projects.

>>>>

In light of the discussion in the philosophy of science, whether paradigm-shifts are tool- or idea-driven, we offer the hypothesis that a tool-driven change of paradigms is underway in the biological sciences and, more subtly, in archaeology. The tool which drives this change is the use of scripting languages in computational approaches of these disciplines. We  show that the use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be referenced more often, giving the authors more visiblity and influence (for some critical aspects on open scholarship see Huggett in this volume). This also is an indicator for the increase in popularity of R.

@@

As with data (see Fitzpatrick in this volume), sharing code in public takes a little time to ensure that it is fit for others to read. Our casual observations of the ways in which archaeologists are using and sharing code show a high degree of variability in code style and organisation, indicating that currently most archaeologists are independantly solving problems of how to write and share code through trial and error, rather than following a model or widely-used approach. This suggests to us that there is some scope to improve the ease and efficiency of code-use in archaeology by studying the exsiting practice and providing a model that others can easily adapt to their own research projects.
